### 2020-04-17 README.Rmd

---
title: "Monitoreo visual"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Documentación