### 2020-10-16 Tablas_Ms_HabitatAves.Rmd
Archivo nuevo con tablas de ocupancia para sp simper

---
title: "Tablas_Ms_ HabitatAves"
author: "Giorgia Graells"
date: "10/16/2020"
output: 
  pdf_document: 
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(tidyverse)
```

```{r, results='asis'}
DF_occu <- tibble( Species= c("Columba livia", "Larus dominicanus", "Leucophaeus pipixcan", "Pelecanus_thagus", "Phalacrocorax_bougainvillii"), 
                   Equation= c("$\\psi_{Cl} = e^{-39.76 + 45.23 \\times native forest - 21.04 \\times gravel + 44.82 \\times sand}$", "$\\psi_{Cl} = e^{12.45}$", "$\\psi_{Cl} = e^{	47.82 -5.09 \\times altitude -231.04 \\times gravel + 5.74 \\times ocean + 	4.40 \\times paved surface}$", "$\\psi_{Cl} = e^{104.91 -0.59 \\times altitude +	6.56\\times native forest 	-2.93 \\times shrublands -18.98\\times sand}$", "$\\psi_{Cl} = e^{-82.05 + 	12.46 \\times crops -15.13 \\times gravel + 	1.38 \\times ocean}$"))

kable(DF_occu, booktabs=T, escape = F) %>% kable_styling(latex_options = "basic")

```

```{r, results='asis'}
DF_occu <- tibble( Species= c("Columba livia", "Larus dominicanus", "Leucophaeus pipixcan", "Pelecanus_thagus", "Phalacrocorax_bougainvillii"), 
                   Equation= c("$\\psi_{Cl} = e^{-39.76 + 45.23 \\times native forest - 21.04 \\times gravel + 44.82 \\times sand}$", "$\\psi_{Cl} = e^{12.45}$", "$\\psi_{Cl} = e^{	47.82 -5.09 \\times altitude -231.04 \\times gravel + 5.74 \\times ocean + 	4.40 \\times paved surface}$", "$\\psi_{Cl} = e^{104.91 -0.59 \\times altitude +	6.56\\times native forest 	-2.93 \\times shrublands -18.98\\times sand}$", "$\\psi_{Cl} = e^{-82.05 + 	12.46 \\times crops -15.13 \\times gravel + 	1.38 \\times ocean}$"))

kable(DF_occu, booktabs=T, escape = F) %>% kable_styling(latex_options = "basic")

```