### 2020-11-21 Tablas_Ms_HabitatAves.Rmd

---
title: "Tablas_Ms_ HabitatAves"
author: "Giorgia Graells"
date: "10/16/2020"
output: 
  pdf_document: 
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
library(tidyverse)
```

```{r,results='asis'}
#Modelo 1= Abundancia ~ Estacion * Especie + AMBIENTE+ (1|Sitio), family = poisson, data =Aves ) #AIC 17716.1
#Modelo 2= Abundancia ~ Especie + AMBIENTE+ (1|Sitio), family = poisson, data =Aves )           #AIC 18241.0
DF_seasson <- tibble(Model= c("$Estacion \\times Especie + Ambiente+ \\frac{1}{Sitio}$", "$Especie + Ambiente + \\frac{1}{Sitio}$"), AIC= c(17716.1, 18241.0))

kable(DF_seasson, booktabs=T, escape = F) %>% kable_styling(latex_options = "basic")

```

```{r, results='asis'}
Inv_DF <-  read.csv("ResumenSimper_Inv2.csv") %>% as.data.frame() 
kable(Inv_DF, booktabs = T)

```

```{r, results='asis'}
Prim_DF <-  read.csv("ResumenSimper_Prim2.csv") %>% as.data.frame()
kable(Prim_DF, booktabs = T)

```

```{r, results='asis'}
DF_occu <- tibble( Species= c("Columba livia", "Larus dominicanus", "Leucophaeus pipixcan", "Pelecanus thagus", "Phalacrocorax bougainvillii"), 
                   Equation= c("$\\psi_{Cl} = e^{-39.76 + 45.23Na - 21.04Gr + 44.82Sa}$",
                               "$\\psi_{Ld} = e^{12.45}$", 
                               "$\\psi_{Lp} = e^{47.82 -5.09Al -231.04Gr + 5.74Oc + 	4.40Pa}$", 
                               "$\\psi_{Pt} = e^{104.91 -0.59Al +	6.56Na	-2.93Sh -18.98Sa}$", 
                               "$\\psi_{Pb} = e^{-82.05 + 12.46Cr -15.13Gr + 	1.38Oc}$"))

kable(DF_occu, booktabs=T, escape = F) %>% kable_styling(latex_options = "basic") %>% column_spec(column = 1, italic = T)

```

```{r, results='asis'}
DF_occu <- tibble( Species= c("Columba livia","Coragyps atratus", "Larosterna inca", "Larus dominicanus", "Leucophaeus modestus", "Pelecanus thagus", "Sephanoides sephaniodes", "Turdus falcklandii"), 
                   Equation= c("$\\psi_{Cl} = e^{-1.77 + 1.83Sa}$", 
                               "$\\psi_{Ca} = e^{7.57 -3.29Pa +	33.80Sa}$", 
                               "$\\psi_{Li} = e^{140.42 -2.38Al	-92.31Na -38.88Sa}$", 
                               "$\\psi_{Ld} = e^{2289.33}$",
                               "$\\psi_{Lm} = e^{190.57 + 2.73Ve -4.07Al -134.07Gr +	20.05Sa}$", 
                               "$\\psi_{Pt} = e^{172.98 -2.01Al	-50.93Na -165.09Gr +	2.47Oc}$", 
                               "$\\psi_{Ss} = e^{-106.91 + 1.64Ve +	7.14Al -43.91Na +	5.09Sh -24.45Tr}$", 
                               "$\\psi_{Tf} = e^{687.42 + 29.20Ve +	48.41Al -34.88Oc	-505.66Gs +	59.49Tr -4.94Pa +	283.63Sa -51.82Tr}$"))

kable(DF_occu, booktabs=T, escape = F) %>% kable_styling(latex_options = "basic") %>% column_spec(column = 1, italic = T)
#dear silvain
```