### 2020-12-10 Ms_HabitatAves.Rmd
Arreglo Ms con presentacion Lab

---
title: "Bird species composition in coastal urban habitats"
output:
  bookdown::word_document2:
    reference_docx: "Para_Template.docx"
bibliography: tesis_ref_2020.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,warning = FALSE,cache = TRUE,comment = FALSE)

library(readxl)
library(tidyverse)
library(vegan)
library(MASS)
library(ggplot2)
library(ggrepel)
library(patchwork)
library(kableExtra)
library(knitr)
library(rasterVis)
library(sf)

```

```{r}
#preparacion de datos registro aves

AvesInv2 <- read_rds("AbundInv_Corregido.rds") 
AvesInv2 <-AvesInv2 %>% ungroup() %>% dplyr::select(-Sitio) %>% mutate_all(~replace(., is.na(.), 0)) %>% sqrt()

AvesPrim2 <- read_rds("AbundPrim_Corregido.rds")
AvesPrim2 <-AvesPrim2 %>% ungroup() %>% dplyr::select(-Sitio)%>% mutate_all(~replace(., is.na(.), 0)) %>% sqrt()

#Datos ambientales  incluidos buffers

Amb<- read_rds("/home/giorgia/Documents/Doctorado tesis/Monitoreo aves/MonitoreoVisualGit/Occdata_occu.rds")
Amb <- Amb %>% 
  dplyr::select(Sitio,CobVeg, AMBIENTE, Distancia_rio, Altura, `Buffer_2200_Bosque Nativo`, `Buffer_2200_Cultivos`, `Buffer_2200_Grava`, `Buffer_2200_Oceano`, `Buffer_2200_Pastizales`, `Buffer_2200_Matorrales`, `Buffer_2200_Sup impermeables`, `Buffer_2200_Suelo arenoso`, `Buffer_2200_Plantación de árboles`) %>% 
  rename( `Vegetation cover` =CobVeg) %>% rename( `Distance to river` =Distancia_rio) %>% 
  rename( `Native forest` =`Buffer_2200_Bosque Nativo`) %>% rename(Crops =`Buffer_2200_Cultivos`) %>% 
  rename( Gravel =`Buffer_2200_Grava`) %>% rename( Ocean =`Buffer_2200_Oceano`) %>% 
  rename( Grassland =`Buffer_2200_Pastizales`) %>% rename( `Shrublands` =`Buffer_2200_Matorrales`) %>% 
  rename( `Paved surface` =`Buffer_2200_Sup impermeables`) %>% rename( `Sand` =`Buffer_2200_Suelo arenoso`) %>% 
  rename( `Tree plantation` =`Buffer_2200_Plantación de árboles`) %>% 
  rename( Altitude = Altura)

rownames(Amb) <- Amb$Sitio

Amb <- Amb %>% dplyr::select(-Sitio) %>%
  mutate(AMBIENTE= str_replace(AMBIENTE, "URBANO", "Urban"),
         AMBIENTE= str_replace(AMBIENTE, "VERDE", "Green area"),
         AMBIENTE= str_replace(AMBIENTE, "ROCA INTERVENIDA", "Modified rocky shore"),
         AMBIENTE= str_replace(AMBIENTE, "PLAYA INTERVENIDA", "Urbanized beach"),
         AMBIENTE= str_replace(AMBIENTE, "ROCA NATURAL", "Natural rocky shore"),
         AMBIENTE= str_replace(AMBIENTE,"PLAYA NATURAL", "Natural beach"))
```

# Introduction

1. Establish importance of the field- provide background, keywords- present the problem

[Establish the importance of this research topic]
During 2018, more than half of world’s population lived in urban settlements with cities growing in both size and number [@UN_WorldCities2018].

[Provide general background information]
Urbanization is a complex process that causes profound changes in the ecological functioning of the landscaped gradually results in a changing spatial structure [@antrop2000changing] being one of the mayor causes of species loss [@czech2000economic].

[Establish the importance of this research topic, more specific/detailed way]
Some of these losses of native species have promoted the replacement by non-native species [@mckinney2006urbanization]. These is because few species can adapt better to urban and suburban conditions [@mckinney1999biotic], affecting at ecosystemic level [grassland: @zeeman2017biotic], biomes [temperate: @zeeman2017biotic], and taxonomic groups [insects: @knop2016biotic, but mainly studies in birds and plants].

[Provide general background information, more specific/detailed way]

Studies suggest that bird communities in cities are more similar to urban communities nearby than to communities in surrounding natural environments. To bird assembles local features are more important than surrounding landscapes @clergeau2001urban,

2. Previous an current research and contributions

[Describe general problem area or the current research focus of the field]

Cities in coastal areas combine many key economic activities such as marine transport, shipping of goods, fishing, recreation, and tourism. These areas are attractive for people to settle and by 2017, about 40%  of the world’s population live within 100 km of the coast [@UNocean_fact2017].
Species and ecosystems of coastal areas are often threatened by the growing urbanization, the landscape change, and also by the vulnerable that this area is naturally

[Provide transition between the general problem area and the literature review]

Even this scenario, urbanization has been mostly focus in terrestrial aspects, and even when more than the half of the cities in the world rise on the coast, there is a lack of studies addressing marine and terrestrial interactions.

3. Locate the gap-describe the problem to address-present a prediction to be tested

[Provide brief overview of key research projects in this area]
Urban homogenization has been studied in cities, focused in terrestrial variables
Considering bird communities in urban areas depend of the peri-urban composition of species, there are no studies in the effects of marine habitats

[Describe gaps in the research]
There is a gap in studies that focus in coastal ecotone and the influence of urban areas
focusing on marine and terrestrial interaction of species in an urbanized area and how birds use this as habitats

[Describe paper itself]
This research aim to identified the use of coastal urban habitats by birds describing the composition of species in different environments, considering bird species origin (marine and terrestrial).
with the aim of: 1. Identify bird species assemble of habitats in coastal urban environments, specifically the interaction between marine and inland birds, 2. identify how urbanization affects those bird assembles, 3. Evaluate iconic bird species habitat occupation, and 4. describe bird behavior in those habitats

4. Describe the present paper
[Give details about the methodology reported in the paper]
The present study was carried out in the Gran Valparaíso, a metropolitan area that includes three municipalities in central Chile

[Announces the findings]
Urban habitats in coastal areas present the conditions that allow specific species composition, cohabiting marine and inland bird species

# Methodology

1. Overview-purpose of the work-give source of material/equipment used

[General overview of this section]

In this research, we conducted bird surveys in six coastal urban environments in the Greater Valparaíso, estimated species composition and analised occupancy of key species in those environments.

[Provide background information and justification]

The Greater Valparaíso is the largest metropolitan coastal area in Central Chile and the second most populated urban area in the country (935,602 people according to @censo2017chile). The three coastal cities: Valparaíso, the mayor port in the country; Viña del Mar, with high touristic infrastructure; and Concón, a small area that grew from a rural beach to a one of place of mayor urban growth in the las txx years.

[Provide an overview of the procedure/method itself]

Bird abundance obtained from the field was corrected by detection capacity of sampling using occupancy models with DiversityOcuppancy package [@divoccu]. To estimate species assembles for habitats, the vegan package [@Vegan] was used. 
Particular species were analized in a continuos environmental gradient using ocupancy probability from occupancy models also with DiversityOcuppancy package [@divoccu].

2. Details about materials and methods- justify choices

[Provide details about what was done and used and shows that care was taken]
Six coastal habitat were defined in terms of its proximity to the coast, level of urbanization and land cover: natural rocky shore, modified rocky shore, natural beach, and urbanized beach, green area, and urban center (Fig. \@ref(fig:Ambientes)). 
Six replicas of theses habitats were sampled by visual bird surveys in south hemisfer winter and spring of 2019. Between four and six surveys were conducted daily randomly during sunrise and the three hours later for 10 minutes in each site. All species that were seen and heard were recorded.

Seasonality was tested with a mixed GLM with abundance as a response variable and a random effect of sample sites. We tested a model where seasonality has a interaction with the habitats defined in this research.

[Continuing to describe what was done in detail, using language which comunicate that care was taken]

Bird commmunities for each habitat were analised using ordination methods with vegan package [@Vegan] in R with species abundances corrected by detection capacity, using occupancy models with DiversityOcuppancy package [@divoccu](see table 1).
We  determined dissimilarity distances among sample sites and tested significant differences among habitat with analysis of similarities (ANOSIM). 
To determine species contribution to the differrence between habitats we used Bray-Curtis dissimilaries with SIMPER analysis.

In order to understand habitat use by birds, environmental variables were prepared from landcover layers.
A 2000m diameter buffer was made

Environmental variables used by bird assembles were tested with a permanova for each season of sampling

```{r Ambientes, fig.cap="The six environments selected to sample: Natural rocky shore, modified rocky shore, natural beach, modified beach, green area, and urban center."}
#figura 1 con los 6 ambientes costeros estudiados

include_graphics("Ambientes_Aves.png")
```

3. Relate to other studies

[Describe what was done by referring to existing mehods in the literature]

To independize our analysis from habitats already defined, ocupancy models of SIMPER selected species were developed and projected over the study area.

[Provide more detailed information about the method and show that was a good choice]
Covariables of detection and occupation are describedd in table 1.
Ocupancy models were proyected to the study area.

4. Indicate where problems occurred

[Mention possible difficulty in the methodology]

```{r}
#Covariables de deteccion y ocupaciono ocupadas para los modelos
```

### Results

1. Revisting the research aim/existing research

Coastal cities in central Chile probed to allow habitats were marine and terrestrial bird species can cohabitat. Urbanization is a big factor in bird assemblages.

2. General overview of results

Species richnes is similar between same natural and modified environments. 
There are distingible birs species assembles where only a few species contribute to the difference between them. For these habitats, bird assemblages differ between winter and spring seasons.
Environmental variables can be responsible of both community identities and use of habitat individualy

3. Invitations to view results

```{r PredRiqueza}

# GLMmixto con variables aleatorias
library(MuMIn)

AvesInv_riq <- readRDS("AvesInv_riq.rds")
AvesPrim_riq <- readRDS("AvesPrim_riq.rds")

nuevoinv <- AvesInv_riq %>% ungroup() %>% mutate(Temporada="Winter") %>% 
  mutate(Ambiente = case_when(AMBIENTE == "PLAYA INTERVENIDA"| AMBIENTE =="PLAYA NATURAL"~"Beach",
                              AMBIENTE =="ROQUERIO INTERVENIDO"| AMBIENTE =="ROQUERIO NATURAL"~"Rocky shore",
                              AMBIENTE =="URBANO" | AMBIENTE == "VERDE"~"Land")) %>% 
  mutate(Grado_modificacion=case_when(AMBIENTE == "PLAYA INTERVENIDA"| AMBIENTE =="ROQUERIO INTERVENIDO" |AMBIENTE =="URBANO"~"Modified",
                                      AMBIENTE == "PLAYA NATURAL"| AMBIENTE =="ROQUERIO NATURAL" |AMBIENTE =="VERDE"~"Natural" ))

nuevoprim <- AvesPrim_riq %>% ungroup() %>% mutate(Temporada="Spring") %>% 
  mutate(Ambiente = case_when(AMBIENTE == "PLAYA INTERVENIDA"| AMBIENTE =="PLAYA NATURAL"~"Beach",
                              AMBIENTE =="ROQUERIO INTERVENIDO"| AMBIENTE =="ROQUERIO NATURAL"~"Rocky shore",
                              AMBIENTE =="URBANO" | AMBIENTE == "VERDE"~"Land")) %>% 
  mutate(Grado_modificacion=case_when(AMBIENTE == "PLAYA INTERVENIDA"| AMBIENTE =="ROQUERIO INTERVENIDO" |AMBIENTE =="URBANO"~"Modified",
                                      AMBIENTE == "PLAYA NATURAL"| AMBIENTE =="ROQUERIO NATURAL" |AMBIENTE =="VERDE"~"Natural" ))

Riqueza <- bind_rows(nuevoinv, nuevoprim)

# Modelo
Modelo <- glm(Riqueza ~ Temporada + Ambiente + Grado_modificacion + Habitat +  Temporada:Ambiente + Ambiente:Grado_modificacion , family = poisson, data =Riqueza ) 
options(na.action = "na.fail")
#Seleccion de modelos

Selected <- dredge(Modelo, m.lim = c(0, floor(nrow(Riqueza)/10)))

Final <- subset(Selected, delta < 2)
Final_conpeso <- model.sel(Final)

Final_DF <- as.data.frame(Final_conpeso)
#kable(Final_DF, digits = 2, booktabs=T) %>%  kable_styling(latex_options = c("scale_down"), font_size = 9)

NewData <- Riqueza %>% dplyr::select(-Riqueza, -Sitio, -AMBIENTE) %>% distinct()

Predicciones <- predict(Modelo, NewData, type = "response",se.fit = T)

NewData$Pred <- Predicciones$fit
NewData$SE <- Predicciones$se.fit

ggplot(NewData, aes(y=Pred, x=Habitat)) + geom_col(aes(fill=Grado_modificacion ), position="dodge") +
  scale_fill_manual(name="Intervention", aesthetics = c("fill", "color"),values = c( '#5ab4ac','#d8b365')) +  theme_bw()+ ylab("Species richness")+
   geom_errorbar(aes(ymin=Pred-SE, ymax=Pred+SE), position = position_dodge2(width = 0.1, padding = 0.7))+
  facet_grid(Ambiente ~ Temporada) + xlab("Habitat of origin")

```

Species richnes is similar between two seasons of sampling, environments, nor habitat of origin of species. Althogh modified anvironmnet have less species for site, only urban centers are distinct.

```{r DataNMDS, results='hide'}
##### METAMDS: Non-metric Multidimensional scaling- NMDS

#########################################################
#Función para sacar los datos y ocupar ggplot
TidyNMDS <- function(MDS, COVS){
  DF <- COVS
  DF2 <- MDS$points %>% as.data.frame()
  DF <- bind_cols(DF2, DF)
  return(DF = DF)
}
#########################################################
#Invierno
Inv_nmds2_bray <- suppressMessages(metaMDS(AvesInv2, autotransform = FALSE, distance="bray", k=2, try=20, trymax=100,  wascores=FALSE, trace=1, plot=FALSE))
#Primavera
Prim_nmds2_bray <-suppressMessages(metaMDS(AvesPrim2,autotransform = FALSE, distance="bray", k=2, try=20, trymax=100, wascores=FALSE, trace=1, plot=FALSE)) 
```

```{r NMDS, fig.width=9, fig.height=4, fig.cap="Sites distribution in A winter and B spring"}
#Funcion para grafico NMDS en ggplot Invierno
Tidy_Inv_nmds2_bray <- TidyNMDS(MDS = Inv_nmds2_bray, COVS = Amb)

Inv <- ggplot(Tidy_Inv_nmds2_bray)+geom_point(aes(x=MDS1, y=MDS2, color=AMBIENTE))+theme_classic() + 
  scale_color_manual(name="Coastal habitats",  values = c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33'))

#Funcion para grafico NMDS en ggplot Primavera
Tidy_Prim_nmds2_bray <- TidyNMDS(MDS = Prim_nmds2_bray, COVS = Amb)

Prim <-ggplot(Tidy_Prim_nmds2_bray)+geom_point(aes(x=MDS1, y=MDS2, color=AMBIENTE))+theme_classic() + 
  scale_color_manual(name="Coastal habitats", values = c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33'))

#Ambos graficos con paquete patchwork
Inv+Prim + plot_annotation(tag_levels = 'A')+ plot_layout(guides = 'collect')
```

Corrected abundances by detection probability

Coastal urban habitats show specific composition of species abundances for winter and spring seassons (Tabla ModGLMEstacion). Distribution of sites in graphical NMDS (Fig. \@ref(fig:NMDS)) showed that both seassons presented differences in abundance of species mostly in the inland sites. Additionally, coastal modified sites, as those in urbanized beach and modified rocky shore habitats, are closer to urban centres than coastal natural sites.

```{r DataAnosim, results='hide'}
## test: ANOSIM

#Invierno
Diss_Inv2 <- vegdist(AvesInv2, method="bray", binary=FALSE, diag=TRUE, upper=TRUE, na.rm = FALSE)
#con base de datos por abundancia
Anosim_Diss_Inv2<-anosim(Diss_Inv2, grouping=Amb$AMBIENTE, permutations = 9999, distance = "bray", strata = NULL, parallel = getOption("mc.cores"))
Summ_AnosimInv <-summary(Anosim_Diss_Inv2)

#Primavera
Diss_Prim2 <- vegdist(AvesPrim2, method="bray", binary=FALSE, diag=TRUE, upper=TRUE, na.rm = FALSE)
#con base de datos por abundancia
Anosim_Diss_Prim2<-anosim(Diss_Prim2, grouping=Amb$AMBIENTE, permutations = 9999, distance = "bray",strata = NULL, parallel = getOption("mc.cores")) 
summary(Anosim_Diss_Prim2)

```

```{r ANOSIM, fig.width=9, fig.cap="Dissimilarity ranking between an within the habitats in A winter (R=0.584, P<0.001) and B spring (R=0.573, P< 0.001)"}
Inv <- ggplot()+ geom_boxplot(aes(x=Anosim_Diss_Inv2$class.vec, y=Anosim_Diss_Inv2$dis.rank), notch = TRUE, varwidth=TRUE)+theme_classic()+
  xlab("Variables")+ ylab("Dissimmilarity rank")

Prim <- ggplot()+ geom_boxplot(aes(Anosim_Diss_Prim2$class.vec, y=Anosim_Diss_Prim2$dis.rank), notch = TRUE, varwidth=TRUE)+theme_classic()+
 xlab("Variables")+ ylab("Dissimmilarity rank")

Inv / Prim +plot_annotation(tag_levels = 'A')
```

Dissimilary rank shows with an 95% confidence interval that dissimilarities withing habitat are lower than between habitats, this demonstrate that group of species for each habitat are different. There is one exception shown by natural beach habitat during winter, here variability of species in the habitat is as high as we consider all habitats together (Fig. \@ref(fig:ANOSIM))

```{r DataPermInv}
#Inv

Mod0 <- adonis2(AvesInv2 ~ 1, data = Amb, by = "terms")
Mod1 <- adonis2(AvesInv2 ~ Altitude +  Ocean +`Vegetation cover`, data = Amb, by = "terms")###################### Menor AIC
Mod2 <- adonis2(AvesInv2 ~ Altitude + `Paved surface`+ Shrublands+`Vegetation cover`, data = Amb, by = "terms")
Mod3 <- adonis2(AvesInv2 ~ Altitude + `Paved surface`+ Ocean , data = Amb, by = "terms")
Mod4 <- adonis2(AvesInv2 ~ Altitude +  Ocean +`Vegetation cover`+ Shrublands, data = Amb, by = "terms")
Mod5 <- adonis2(AvesInv2 ~ Altitude + `Paved surface`+ Shrublands+ Ocean, data = Amb, by = "terms")

AICc.PERMANOVA2 <- function(adonis2.model) {

  # check to see if object is an adonis2 model...

  if (is.na(adonis2.model$SumOfSqs[1]))
    stop("object not output of adonis2 {vegan} ")

  RSS <- adonis2.model$SumOfSqs[ length(adonis2.model$SumOfSqs) - 1 ]
  MSE <- RSS / adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  nn <- adonis2.model$Df[ length(adonis2.model$Df) ] + 1

  k <- nn - adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  # AIC : 2*k + n*ln(RSS/n)
  # AICc: AIC + [2k(k+1)]/(n-k-1)

  AIC <- 2*k + nn*log(RSS/nn)
  AICc <- AIC + (2*k*(k + 1))/(nn - k - 1)

  output <- data.frame(AICc = AICc, k = k, N = nn)

  return(output)

}

DF<- data.frame(Modelo = paste0("Mod", 0:5), AICc = NA)

DF$AICc[1] <- AICc.PERMANOVA2(Mod0)$AICc
DF$AICc[2] <-  AICc.PERMANOVA2(Mod1)$AICc
DF$AICc[3] <- AICc.PERMANOVA2(Mod2)$AICc
DF$AICc[4] <-  AICc.PERMANOVA2(Mod3)$AICc
DF$AICc[5] <-  AICc.PERMANOVA2(Mod4)$AICc
DF$AICc[6] <-  AICc.PERMANOVA2(Mod5)$AICc

```

```{r DataPermPrim}
#Prim

Mod0 <- adonis2(AvesPrim2 ~ 1, data = Amb, by = "terms")
Mod1 <- adonis2(AvesPrim2 ~ Altitude +  Ocean +`Vegetation cover`, data = Amb, by = "terms")
Mod2 <- adonis2(AvesPrim2 ~ Altitude + `Paved surface`+ Shrublands+`Vegetation cover`, data = Amb, by = "terms")
Mod3 <- adonis2(AvesPrim2 ~ Altitude + `Paved surface`+ Ocean , data = Amb, by = "terms")########################     Menor AIC
Mod4 <- adonis2(AvesPrim2 ~ Altitude +  Ocean +`Vegetation cover`+ Shrublands, data = Amb, by = "terms")
Mod5 <- adonis2(AvesPrim2 ~ Altitude + `Paved surface`+ Shrublands+ Ocean, data = Amb, by = "terms")

AICc.PERMANOVA2 <- function(adonis2.model) {

  # check to see if object is an adonis2 model...

  if (is.na(adonis2.model$SumOfSqs[1]))
    stop("object not output of adonis2 {vegan} ")

  RSS <- adonis2.model$SumOfSqs[ length(adonis2.model$SumOfSqs) - 1 ]
  MSE <- RSS / adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  nn <- adonis2.model$Df[ length(adonis2.model$Df) ] + 1

  k <- nn - adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  # AIC : 2*k + n*ln(RSS/n)
  # AICc: AIC + [2k(k+1)]/(n-k-1)

   AIC <- 2*k + nn*log(RSS/nn)
  AICc <- AIC + (2*k*(k + 1))/(nn - k - 1)

  output <- data.frame(AICc = AICc, k = k, N = nn)

  return(output)

}

DF<- data.frame(Modelo = paste0("Mod", 0:5), AICc = NA)

DF$AICc[1] <- AICc.PERMANOVA2(Mod0)$AICc
DF$AICc[2] <-  AICc.PERMANOVA2(Mod1)$AICc
DF$AICc[3] <- AICc.PERMANOVA2(Mod2)$AICc
DF$AICc[4] <-  AICc.PERMANOVA2(Mod3)$AICc
DF$AICc[5] <-  AICc.PERMANOVA2(Mod4)$AICc
DF$AICc[6] <-  AICc.PERMANOVA2(Mod5)$AICc

```

Altitud, oceano y cob veg en invierno
Altitud, oceano y sup imp en primavera

4. Specific/key results in details

Both sampled seasons showed that few species contribute to differences between habitats. Acccording to simper analysis, during winter species were *Pelecanus thagus*, *Larus dominicanus*, *Coragyps atratus*, *Larosterna inca*, *Leucophaeus modestus*, *Columba livia*, *Turdus falcklandii*, and *Sephanoides sephaniodes*; and during spring, *Leucophaeus pipixcan*, *Larus dominicanus*, *Columba livia*, *Larosterna inca*, *Phalacrocorax bougainvillii*, and *Pelecanus thagus*.  (For a complete species list of SIMPER, see Suplementary material SimperPrim, SimperInv).

```{r OccuInv, fig.width=10, fig.height=7}

library(sf)
library(raster)
library(ggrepel)
Pelecanus <-read_rds("PredInv_Pelecanus_thagus.rds")
Larus <-read_rds("PredInv_Larus_dominicanus.rds") 
Coragyps <-read_rds("PredInv_Coragyps_atratus.rds")
Larosterna <-read_rds("PredInv_Larosterna_inca.rds")
Leucophaeus <-read_rds("PredInv_Leucophaeus_modestus.rds")
Columba <-read_rds("PredInv_Columba_livia.rds")
Turdus <-read_rds("PredInv_Turdus_falcklandii.rds")
Sephanoides <-read_rds("PredInv_Sephanoides_sephaniodes.rds")

Nombres <- c("Pelecanus", "Larus", "Coragyps", "Larosterna", "Leucophaeus", "Columba", "Turdus", "Sephanoides")
Species <- list(Pelecanus, Larus, Coragyps, Larosterna, Leucophaeus, Columba, Turdus, Sephanoides)

temp <- list()

for(i in 1:length(Species)){
  temp[[i]] <- Species[[i]] %>% as("SpatialPixelsDataFrame") %>% as.data.frame()
  colnames(temp[[i]])[1] <- Nombres[i]
}

Species_DF <- temp %>% reduce(full_join) %>% dplyr::relocate(Pelecanus, .after = y) %>% pivot_longer(Pelecanus:Sephanoides, names_to = "Species", values_to = "Occupancy")

Sitios <- read.csv("/home/giorgia/Documents/Doctorado tesis/Monitoreo aves/MonitoreoVisualGit/Muestreo aves jun-jul 2019/Monitoreo_punto/Reg_completo_aves_inv.csv") %>% 
  dplyr::filter(Sitio %in% c("TERMINAL", "PLAYA LA BOCA", "PARQUE EL LITRE", "CALETA EL MEMBRILLO", "ROCA OCEANICA","SECTOR 5")) %>% 
  mutate(Sitio= str_replace(Sitio, "TERMINAL", "T"),
         Sitio= str_replace(Sitio, "PLAYA LA BOCA", "PLB"),
         Sitio= str_replace(Sitio, "PARQUE EL LITRE", "PEL"),
         Sitio= str_replace(Sitio, "CALETA EL MEMBRILLO", "CEM"),
         Sitio= str_replace(Sitio, "ROCA OCEANICA", "RO"),
         Sitio= str_replace(Sitio, "SECTOR 5", "PR")) %>% 
  dplyr::select(Sitio, Latitud, Longitud) %>% distinct() %>% st_as_sf(coords=c("Longitud","Latitud"),crs="+proj=longlat +datum=WGS84 +no_defs") %>% 
  st_transform(crs="+proj=utm +zone=19 +south +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")

OcuI <- ggplot() + geom_raster(data = Species_DF, aes(x = x, y = y, fill = Occupancy)) + geom_sf(data = Sitios, color="red") +
    ggrepel::geom_text_repel(data = Sitios,  aes(label = Sitio,  geometry = geometry),  stat = "sf_coordinates", force = 1, 
nudge_x=-5000, nudge_y = 1000)+ 
  theme_classic() + scale_fill_viridis_c("Occupancy", option = "plasma", direction = -1) + labs(x = NULL, y = NULL) + facet_wrap(~Species, nrow=1)+ theme(legend.position = "none")

```

```{r OccuPrim, fig.width=8, fig.height=7}
Columba <-read_rds("PredPrim_Columba_livia.rds")
Larosterna <-read_rds("PredPrim_Larosterna_inca.rds")
Larus <-read_rds("PredPrim_Larus_dominicanus.rds")
Leucophaeus <-read_rds("PredPrim_Leucophaeus_pipixcan.rds")
Phalacrocorax <-read_rds("PredPrim_Phalacrocorax_bougainvillii.rds")
Pelecanus <-read_rds("PredPrim_Pelecanus_thagus.rds")

Nombres <- c("Columba", "Larosterna", "Larus","Leucophaeus", "Phalacrocorax", "Pelecanus")
Species <- list(Columba, Larosterna, Larus, Leucophaeus, Phalacrocorax, Pelecanus)

temp <- list()

for(i in 1:length(Species)){
  temp[[i]] <- Species[[i]] %>% as("SpatialPixelsDataFrame") %>% as.data.frame()
  colnames(temp[[i]])[1] <- Nombres[i]
}

Species_DF <- temp %>% reduce(full_join) %>% dplyr::relocate(Columba, .after = y) %>% pivot_longer(Columba:Pelecanus, names_to = "Species", values_to = "Occupancy")

Sitios <- read.csv("/home/giorgia/Documents/Doctorado tesis/Monitoreo aves/MonitoreoVisualGit/Muestreo aves jun-jul 2019/Monitoreo_punto/Reg_completo_aves_inv.csv") %>% 
 dplyr::filter(Sitio %in% c("TERMINAL", "PLAYA LA BOCA", "PARQUE EL LITRE", "CALETA EL MEMBRILLO", "ROCA OCEANICA","SECTOR 5")) %>% 
  mutate(Sitio= str_replace(Sitio, "TERMINAL", "T"),
         Sitio= str_replace(Sitio, "PLAYA LA BOCA", "PLB"),
         Sitio= str_replace(Sitio, "PARQUE EL LITRE", "PEL"),
         Sitio= str_replace(Sitio, "CALETA EL MEMBRILLO", "CEM"),
         Sitio= str_replace(Sitio, "ROCA OCEANICA", "RO"),
         Sitio= str_replace(Sitio, "SECTOR 5", "PR")) %>% 
  dplyr::select(Sitio, Latitud, Longitud) %>% distinct() %>% st_as_sf(coords=c("Longitud","Latitud"),crs="+proj=longlat +datum=WGS84 +no_defs") %>% 
  st_transform(crs="+proj=utm +zone=19 +south +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")

OcuP <- ggplot() + geom_raster(data = Species_DF, aes(x = x, y = y, fill = Occupancy)) + geom_sf(data = Sitios, color="red") +
    ggrepel::geom_text_repel(data = Sitios,  aes(label = Sitio,  geometry = geometry),  stat = "sf_coordinates", force = 1, 
nudge_x=-5000, nudge_y = 1000)+ 
  theme_classic() + scale_fill_viridis_c("Occupancy", option = "plasma", direction = -1) + labs(x = NULL, y = NULL) + facet_wrap(~Species, nrow=1)

```

```{r Ocupancia, fig.width=13, fig.height=6}

layout <- "
AAAA
BBB#
"
OcuI + OcuP+plot_annotation(tag_levels = 'A')+ plot_layout(design = layout)

```

Predictions in the studied area indicates differences of habitat use (Fig. \@ref(fig:OccuInv) and \@ref(fig:OccuPrim))

Species such as xxx, showed differernces in habitat use between seasons.
(Fig. \@ref(fig:OccuInv)), (Fig. \@ref(fig:OccuPrim))

During winter, community assemblages were explain by three environmental variables: coastal urban habitats, crops and tree plantations. 
In spring, variables were coastal urban habitats and native forest.

species for each habitat
Occupancy model map show

```{r}
#Resultados occupancia para especies dde Simper
ModelosOccuPrim <- read_rds("ModelosOccuPrim_Ms.rds")
#Para generar las ecuaciones **Resultado_OccuPrim_Ms.rds** y **Resultado_OccuInv_Ms.rds**
```

Finally, use of these habitats can be seen with the activity record of bird species.

```{r Comportamiento, fig.height=8, fig.width=10}
# ComportamientoInv <- read_rds("ComportamientoInv.rds")
# I <-ggplot(ComportamientoInv) + 
#   geom_col(aes(x=AMBIENTE, y=Total_ind, fill=Comportamiento), color="black") + theme_classic()+ facet_wrap(~Cientifico, ncol=4, scales = "free_y")+  
#   theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))+ ylab("Número individuos") +xlab(" ")+
#   scale_fill_manual(aesthetics = "fill", values = c('#a6cee3','#1f78b4','#33a02c','#e31a1c','#fdbf6f','#ff7f00'), name="Bird behavior")
# 
# ComportamientoPrim <- read_rds( "ComportamientoPrim.rds")
# 
# P <-ggplot(ComportamientoPrim) + 
#   geom_col(aes(x=AMBIENTE, y=Total_ind, fill=Comportamiento), color="black") + theme_classic()+ facet_wrap(~Cientifico, ncol=3, scales = "free_y")+  
#   theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 1))+ ylab("Número individuos") +xlab(" ")+
#   scale_fill_manual(aesthetics = "fill", values = c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00'), name="Bird behavior")
# 
# layout <- "
# AAAAAAAA
# BBBBBBB#
# "
# I + P + plot_layout(design = layout) +plot_annotation(tag_levels = 'A')
#

```

Most of the species were detected during the fliying, however some were founded in particular activities.

5. Comparisons with results in other research

Occupancy models for specific species

6. problems with results

7. Possible implications with results

### Discussion/Conclusion

1. Revising previous research- summaries
[Revisit previous research]
[Revisit Introduction to recall specific weakness in methodology used in previous studies]

2. Mapping 
[Revisit methodology used in this study]
[Revisit and summaries the results]
[Show where and how the present work fits into the research "map" in this field]

3. Achievements/contributions- refining implications
[Recall an aspect of the results that represents a positive achievement or contribution of this work]
[Focus in the meaning and implications of the achievements in this work]
[Note if one of the achievement or contributions of the work is a novelty]
[Describe implications of the results, including possible applications]

4. Limitations-current and future work-applications
[Describe limitations which should direct future research]
[Suggest a specific area to be addressed in future work]

### References