### 2021-01-12 Ms_HabitatAves.Rmd
Arreglo intro y resultados

---
title: "Bird species composition in coastal urban habitats"
output:
  bookdown::word_document2:
    reference_docx: "Para_Template.docx"
bibliography: tesis_ref_2020.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,	message = FALSE,warning = FALSE,cache = TRUE,comment = FALSE)

library(readxl)
library(tidyverse)
library(vegan)
library(MASS)
library(ggplot2)
library(ggrepel)
library(patchwork)
library(kableExtra)
library(knitr)
library(rasterVis)
library(sf)

```

```{r}
#preparacion de datos registro aves

AvesInv2 <- read_rds("AbundInv_Corregido.rds") 
AvesInv2 <-AvesInv2 %>% ungroup() %>% dplyr::select(-Sitio) %>% mutate_all(~replace(., is.na(.), 0)) %>% sqrt()

AvesPrim2 <- read_rds("AbundPrim_Corregido.rds")
AvesPrim2 <-AvesPrim2 %>% ungroup() %>% dplyr::select(-Sitio)%>% mutate_all(~replace(., is.na(.), 0)) %>% sqrt()

#Datos ambientales  incluidos buffers

Amb<- read_rds("/home/giorgia/Documents/Doctorado tesis/Monitoreo aves/MonitoreoVisualGit/Occdata_occu.rds")
Amb <- Amb %>% 
  dplyr::select(Sitio,CobVeg, AMBIENTE, Distancia_rio, Altura, `Buffer_2200_Bosque Nativo`, `Buffer_2200_Cultivos`, `Buffer_2200_Grava`, `Buffer_2200_Oceano`, `Buffer_2200_Pastizales`, `Buffer_2200_Matorrales`, `Buffer_2200_Sup impermeables`, `Buffer_2200_Suelo arenoso`, `Buffer_2200_Plantación de árboles`) %>% 
  rename( `Vegetation cover` =CobVeg) %>% rename( `Distance to river` =Distancia_rio) %>% 
  rename( `Native forest` =`Buffer_2200_Bosque Nativo`) %>% rename(Crops =`Buffer_2200_Cultivos`) %>% 
  rename( Gravel =`Buffer_2200_Grava`) %>% rename( Ocean =`Buffer_2200_Oceano`) %>% 
  rename( Grassland =`Buffer_2200_Pastizales`) %>% rename( `Shrublands` =`Buffer_2200_Matorrales`) %>% 
  rename( `Paved surface` =`Buffer_2200_Sup impermeables`) %>% rename( `Sand` =`Buffer_2200_Suelo arenoso`) %>% 
  rename( `Tree plantation` =`Buffer_2200_Plantación de árboles`) %>% 
  rename( Altitude = Altura)

rownames(Amb) <- Amb$Sitio

Amb <- Amb %>% dplyr::select(-Sitio) %>%
  mutate(AMBIENTE= str_replace(AMBIENTE, "URBANO", "Urban"),
         AMBIENTE= str_replace(AMBIENTE, "VERDE", "Green area"),
         AMBIENTE= str_replace(AMBIENTE, "ROCA INTERVENIDA", "Modified rocky shore"),
         AMBIENTE= str_replace(AMBIENTE, "PLAYA INTERVENIDA", "Urbanized beach"),
         AMBIENTE= str_replace(AMBIENTE, "ROCA NATURAL", "Natural rocky shore"),
         AMBIENTE= str_replace(AMBIENTE,"PLAYA NATURAL", "Natural beach"))
```

# Abstract

Six environments present specific assembles and they are determined mostly by variables such as altitude and percentage of the ocean in the areas. From the species ocurrence in those, there are strictly marine species and some can move to inland environments. Most of the terrestrial species are less strict and they manage to get closer to the ocean, specifically in areas with native vegetation such as natural beaches and natural rocky shore (with hay percentage of shrubs and native forest).
Elements of urbanization and closeness to the coast are important factors on bird assemblages

# Introduction

During 2018, more than half of world’s population lived in urban settlements with cities growing in both size and number [@UN_WorldCities2018].
Urbanization is a complex process and one of the major drivers of natural habitat loss and degradation [@grimm2008global] and it has generated profound changes in the ecological functioning of the landscape [@antrop2000changing] being 
Although the impact of urbanisation varies with the speed and intensity of environmental alterations, it is one of the mayor causes of biodiversity loss [@czech2000economic]. (McKinney 2002; Clergeauet al.2006; Aronsonet al.2014).
It has been estudied that loss of native species have promoted the replacement by non-native species [@mckinney2006urbanization]. These is because few species can adapt better to urban and suburban conditions [@mckinney1999biotic], affecting at ecosystemic level [grassland: @zeeman2017biotic], biomes [temperate: @zeeman2017biotic], and taxonomic groups [insects: @knop2016biotic, but mainly studies in birds and plants].
Studies suggest that bird communities in cities are more similar to urban communities nearby than to communities in surrounding natural environments. To bird assembles, local features are more important than surrounding landscapes @clergeau2001urban.

Cities in coastal areas combine many key economic activities such as marine transport, shipping of goods, fishing, recreation, and tourism. These areas are attractive for people to settle and by 2017, about 40%  of the world’s population live within 100 km of the coast [@UNocean_fact2017].
Hence, only in 2009, 67% of magacities corresponded to coastal cities *von glasgow 2013*.
Species and ecosystems of coastal areas are often threatened by the growing urbanization, the landscape change, and also by the vulnerable that this area is naturally
Even this scenario, urbanization has been mostly focus in terrestrial aspects, and even when more than the half of the cities in the world rise on the coast, there is a lack of studies addressing marine and terrestrial interactions.
marine and inland interactions could be importatn

Birds have been largely studied in urban ecology and they are considered one of the most recognised ways to interact with nature [@miller2002conservation; @jones2011appetite]
Urban homogenization has been studied in cities and there is many information about bird species.
Considering bird communities in urban areas depend of the peri-urban composition of species, there are no studies in the effects of marine habitats.
Although birds are the most research species in coastal urban studies, most of the literature have been centered in pollution and human impacts and they have focused on terrestrial ecosystems [@graells2020review].
There is a gap in studies that focus in coastal ecotone and the influence of urban areas focusing on marine and terrestrial interaction of species in an urbanized area and how birds use it.

This research aim to identified the use of coastal urban habitats by birds describing the composition of species in different environments, considering bird species origin (marine and terrestrial).
Specifically, the objectives of this study are: 
1. Identify and characterize bird species assemble in coastal urban environments, specifically the interaction between marine and inland birds, and 
2. identify patterns in spatial use in coastal cities from iconic bird species habitat occupation.

We studied coastal urban diversity of birds in coastal metropolitan area in South America, determined assembles in different environments, relationships with environmental conditions and occupancy of key bird species.
Bird assembles showed to be different among environments, showing that variability of urban habitats in coastal areas allow specific species composition, hence the coexistence of marine and inland bird species.

**Conclusion: La variabilidad ambiental dentro de las ciudades costeras permite la co-ocurrencia de especies marinas y terrestres, determinando ensambles especfíficos para los distintos ambientes**

# Methodology

In this research, we conducted bird surveys in six coastal urban environments in the Greater Valparaíso, estimated species composition and analised occupancy of key species in those environments.
The Greater Valparaíso is the largest metropolitan coastal area in Chile and the second most populated urban area in the country (935,602 people according to @censo2017chile). It is composed by three coastal cities: Valparaíso, the mayor port in the country; Viña del Mar, with high touristic infrastructure; and Concón, a small area that grew from a rural beach to a one of place of mayor urban growth in the las xx years.

```{r}

```

To estimate species assembles for habitats, the vegan package [@Vegan] was used. 
Particular species were analized in a continuos environmental gradient using ocupancy probability from occupancy models also with DiversityOcuppancy package [@divoccu].

Six coastal habitat were defined in terms of its proximity to the coast, level of urbanization and land cover: natural rocky shore, modified rocky shore, natural beach, and urbanized beach, green area, and urban center (Fig. \@ref(fig:Ambientes)). 
Six replicas of theses habitats were sampled by visual bird surveys in south hemisfer winter and spring of 2019. Between four and six surveys were conducted daily and randomly during sunrise and the three hours later for 10 minutes in each site. All species that were seen and heard were recorded in a tablet.

## Data analysis

Bird abundance obtained from the field was corrected by detection capacity of sampling using occupancy models with DiversityOcuppancy package [@divoccu]. 
Seasonality was tested with a mixed GLM with abundance as a response variable and a random effect of sample sites. We tested a model where seasonality has a interaction with the habitats defined in this research.

Bird commmunities for each habitat were analised using ordination methods with vegan package [@Vegan] in R with species abundances corrected by detection capacity, using occupancy models with DiversityOcuppancy package [@divoccu](see table 1).
We  determined dissimilarity distances among sample sites and tested significant differences among habitat with analysis of similarities (ANOSIM). 
To determine species contribution to the difference between habitats we used Bray-Curtis dissimilaries with SIMPER analysis.
Environmental variables used by bird assembles were tested with a permanova for each season of sampling.

```{r Ambientes, fig.cap="The six environments selected to sample: Natural rocky shore, modified rocky shore, natural beach, modified beach, green area, and urban center."}
#figura 1 con los 6 ambientes costeros estudiados

include_graphics("Ambientes_Aves.png")
```

SIMPER selected species ocurrence was analized using ocupancy models with the package DiversityOccupancy in R [@divoccu] . In order to do this, environmental variables were prepared from landcover layers.
For each of the 36 samplig points a 2000m diameter buffer was made using the pecentaje of covering area in the selection.
Covariables of occupation are described in table 1. 
To select the best model of ocurrence, Akaike Information Criteria (AIC) was used. 
Models selected were proyected to the study area showing the probability of use of habitat for each species in the area.

```{r}
#Covariables de deteccion y ocupaciono ocupadas para los modelos
```

### Results

Coastal cities in central Chile probed to allow habitats were marine and terrestrial bird species can cohabitat. *Elements of urbanization and closeness to the coast are important factors on bird assemblages.*

A total of 58 bird species were found in the six coastal urban environments sampled. From those, 37 were present in both seasons.
Species richnes are similar between winter and spring (45 and 48 species, respectively).

Predicted species richness are also similar between same natural and modified environments. Only terrestrial environments showed significative differences, where urban environments showed fewer species that green areas.

```{r PredRiqueza}

# GLMmixto con variables aleatorias
library(MuMIn)

AvesInv_riq <- readRDS("AvesInv_riq.rds")
AvesPrim_riq <- readRDS("AvesPrim_riq.rds")

nuevoinv <- AvesInv_riq %>% ungroup() %>% mutate(Temporada="Winter") %>% 
  mutate(Ambiente = case_when(AMBIENTE == "PLAYA INTERVENIDA"| AMBIENTE =="PLAYA NATURAL"~"Beach",
                              AMBIENTE =="ROQUERIO INTERVENIDO"| AMBIENTE =="ROQUERIO NATURAL"~"Rocky shore",
                              AMBIENTE =="URBANO" | AMBIENTE == "VERDE"~"Land")) %>% 
  mutate(Grado_modificacion=case_when(AMBIENTE == "PLAYA INTERVENIDA"| AMBIENTE =="ROQUERIO INTERVENIDO" |AMBIENTE =="URBANO"~"Modified",
                                      AMBIENTE == "PLAYA NATURAL"| AMBIENTE =="ROQUERIO NATURAL" |AMBIENTE =="VERDE"~"Natural" ))

nuevoprim <- AvesPrim_riq %>% ungroup() %>% mutate(Temporada="Spring") %>% 
  mutate(Ambiente = case_when(AMBIENTE == "PLAYA INTERVENIDA"| AMBIENTE =="PLAYA NATURAL"~"Beach",
                              AMBIENTE =="ROQUERIO INTERVENIDO"| AMBIENTE =="ROQUERIO NATURAL"~"Rocky shore",
                              AMBIENTE =="URBANO" | AMBIENTE == "VERDE"~"Land")) %>% 
  mutate(Grado_modificacion=case_when(AMBIENTE == "PLAYA INTERVENIDA"| AMBIENTE =="ROQUERIO INTERVENIDO" |AMBIENTE =="URBANO"~"Modified",
                                      AMBIENTE == "PLAYA NATURAL"| AMBIENTE =="ROQUERIO NATURAL" |AMBIENTE =="VERDE"~"Natural" ))

Riqueza <- bind_rows(nuevoinv, nuevoprim)

# Modelo
Modelo <- glm(Riqueza ~ Temporada + Ambiente + Grado_modificacion + Habitat +  Temporada:Ambiente + Ambiente:Grado_modificacion , family = poisson, data =Riqueza ) 
options(na.action = "na.fail")
#Seleccion de modelos

Selected <- dredge(Modelo, m.lim = c(0, floor(nrow(Riqueza)/10)))

Final <- subset(Selected, delta < 2)
Final_conpeso <- model.sel(Final)

Final_DF <- as.data.frame(Final_conpeso)
#kable(Final_DF, digits = 2, booktabs=T) %>%  kable_styling(latex_options = c("scale_down"), font_size = 9)

NewData <- Riqueza %>% dplyr::select(-Riqueza, -Sitio, -AMBIENTE) %>% distinct()

Predicciones <- predict(Modelo, NewData, type = "response",se.fit = T)

NewData$Pred <- Predicciones$fit
NewData$SE <- Predicciones$se.fit

ggplot(NewData, aes(y=Pred, x=Habitat)) + geom_col(aes(fill=Grado_modificacion ), position="dodge") +
  scale_fill_manual(name="Intervention", aesthetics = c("fill", "color"),values = c( '#5ab4ac','#d8b365')) +  theme_bw()+ ylab("Species richness")+
   geom_errorbar(aes(ymin=Pred-SE, ymax=Pred+SE), position = position_dodge2(width = 0.1, padding = 0.7))+
  facet_grid(Ambiente ~ Temporada) + xlab("Habitat of origin")

```

```{r DataNMDS, results='hide'}
##### METAMDS: Non-metric Multidimensional scaling- NMDS

#########################################################
#Función para sacar los datos y ocupar ggplot
TidyNMDS <- function(MDS, COVS){
  DF <- COVS
  DF2 <- MDS$points %>% as.data.frame()
  DF <- bind_cols(DF2, DF)
  return(DF = DF)
}
#########################################################
#Invierno
Inv_nmds2_bray <- suppressMessages(metaMDS(AvesInv2, autotransform = FALSE, distance="bray", k=2, try=20, trymax=100,  wascores=FALSE, trace=1, plot=FALSE))
#Primavera
Prim_nmds2_bray <-suppressMessages(metaMDS(AvesPrim2,autotransform = FALSE, distance="bray", k=2, try=20, trymax=100, wascores=FALSE, trace=1, plot=FALSE)) 
```

```{r NMDS, fig.width=9, fig.height=4, fig.cap="MDS sites distribution according to bird asembles in A: winter and B: spring"}
#Funcion para grafico NMDS en ggplot Invierno
Tidy_Inv_nmds2_bray <- TidyNMDS(MDS = Inv_nmds2_bray, COVS = Amb)

Inv <- ggplot(Tidy_Inv_nmds2_bray)+geom_point(aes(x=MDS1, y=MDS2, color=AMBIENTE))+theme_classic() + 
  scale_color_manual(name="Coastal habitats",  values = c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33'))

#Funcion para grafico NMDS en ggplot Primavera
Tidy_Prim_nmds2_bray <- TidyNMDS(MDS = Prim_nmds2_bray, COVS = Amb)

Prim <-ggplot(Tidy_Prim_nmds2_bray)+geom_point(aes(x=MDS1, y=MDS2, color=AMBIENTE))+theme_classic() + 
  scale_color_manual(name="Coastal habitats", values = c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33'))

#Ambos graficos con paquete patchwork
Inv+Prim + plot_annotation(tag_levels = 'A')+ plot_layout(guides = 'collect')
```

Coastal urban habitats show specific composition of species abundances for winter and spring seasons (Tabla ModGLMEstacion). Distribution of sites in graphical MDS (Fig. \@ref(fig:NMDS)) showed that both seassons presented differences in abundance of species mostly in the inland sites. Additionally, coastal modified sites, such as urbanized beach and modified rocky shore, are more similar to urban centres than coastal natural sites.

```{r DataAnosim, results='hide'}
## test: ANOSIM

#Invierno
Diss_Inv2 <- vegdist(AvesInv2, method="bray", binary=FALSE, diag=TRUE, upper=TRUE, na.rm = FALSE)
#con base de datos por abundancia
Anosim_Diss_Inv2<-anosim(Diss_Inv2, grouping=Amb$AMBIENTE, permutations = 9999, distance = "bray", strata = NULL, parallel = getOption("mc.cores"))
Summ_AnosimInv <-summary(Anosim_Diss_Inv2)

#Primavera
Diss_Prim2 <- vegdist(AvesPrim2, method="bray", binary=FALSE, diag=TRUE, upper=TRUE, na.rm = FALSE)
#con base de datos por abundancia
Anosim_Diss_Prim2<-anosim(Diss_Prim2, grouping=Amb$AMBIENTE, permutations = 9999, distance = "bray",strata = NULL, parallel = getOption("mc.cores")) 
summary(Anosim_Diss_Prim2)

```

```{r ANOSIM, fig.width=9, fig.cap="Dissimilarity ranking between an within the habitats in A: winter (R=0.584, P<0.001) and B: spring (R=0.573, P< 0.001)"}
Inv <- ggplot()+ geom_boxplot(aes(x=Anosim_Diss_Inv2$class.vec, y=Anosim_Diss_Inv2$dis.rank), notch = TRUE, varwidth=TRUE)+theme_classic()+
  xlab("Variables")+ ylab("Dissimmilarity rank")

Prim <- ggplot()+ geom_boxplot(aes(Anosim_Diss_Prim2$class.vec, y=Anosim_Diss_Prim2$dis.rank), notch = TRUE, varwidth=TRUE)+theme_classic()+
 xlab("Variables")+ ylab("Dissimmilarity rank")

Inv / Prim +plot_annotation(tag_levels = 'A')
```

Seasonal ANOSIM analysis shows that assembles in each environment are different among them for both seasons studied (R=0.584, P< 0.01 in winter and R=0.573, P< 0.01 in spring).
Dissimilary rank indicates with an 95% confidence interval (notch in boxplot), that dissimilarities withing each habitat are lower than between them. 
There is one exception shown by natural beach habitat during winter, here variability of species in the habitat is as high as we consider all habitats together (Fig. \@ref(fig:ANOSIM)).

```{r DataPermInv}
#Inv

Mod0 <- adonis2(AvesInv2 ~ 1, data = Amb, by = "terms")
Mod1 <- adonis2(AvesInv2 ~ Altitude +  Ocean +`Vegetation cover`, data = Amb, by = "terms")###################### Menor AIC
Mod2 <- adonis2(AvesInv2 ~ Altitude + `Paved surface`+ Shrublands+`Vegetation cover`, data = Amb, by = "terms")
Mod3 <- adonis2(AvesInv2 ~ Altitude + `Paved surface`+ Ocean , data = Amb, by = "terms")
Mod4 <- adonis2(AvesInv2 ~ Altitude +  Ocean +`Vegetation cover`+ Shrublands, data = Amb, by = "terms")
Mod5 <- adonis2(AvesInv2 ~ Altitude + `Paved surface`+ Shrublands+ Ocean, data = Amb, by = "terms")

AICc.PERMANOVA2 <- function(adonis2.model) {

  # check to see if object is an adonis2 model...

  if (is.na(adonis2.model$SumOfSqs[1]))
    stop("object not output of adonis2 {vegan} ")

  RSS <- adonis2.model$SumOfSqs[ length(adonis2.model$SumOfSqs) - 1 ]
  MSE <- RSS / adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  nn <- adonis2.model$Df[ length(adonis2.model$Df) ] + 1

  k <- nn - adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  # AIC : 2*k + n*ln(RSS/n)
  # AICc: AIC + [2k(k+1)]/(n-k-1)

  AIC <- 2*k + nn*log(RSS/nn)
  AICc <- AIC + (2*k*(k + 1))/(nn - k - 1)

  output <- data.frame(AICc = AICc, k = k, N = nn)

  return(output)

}

DF<- data.frame(Modelo = paste0("Mod", 0:5), AICc = NA)

DF$AICc[1] <- AICc.PERMANOVA2(Mod0)$AICc
DF$AICc[2] <-  AICc.PERMANOVA2(Mod1)$AICc
DF$AICc[3] <- AICc.PERMANOVA2(Mod2)$AICc
DF$AICc[4] <-  AICc.PERMANOVA2(Mod3)$AICc
DF$AICc[5] <-  AICc.PERMANOVA2(Mod4)$AICc
DF$AICc[6] <-  AICc.PERMANOVA2(Mod5)$AICc

```

```{r DataPermPrim}
#Prim

Mod0 <- adonis2(AvesPrim2 ~ 1, data = Amb, by = "terms")
Mod1 <- adonis2(AvesPrim2 ~ Altitude +  Ocean +`Vegetation cover`, data = Amb, by = "terms")
Mod2 <- adonis2(AvesPrim2 ~ Altitude + `Paved surface`+ Shrublands+`Vegetation cover`, data = Amb, by = "terms")
Mod3 <- adonis2(AvesPrim2 ~ Altitude + `Paved surface`+ Ocean , data = Amb, by = "terms")########################     Menor AIC
Mod4 <- adonis2(AvesPrim2 ~ Altitude +  Ocean +`Vegetation cover`+ Shrublands, data = Amb, by = "terms")
Mod5 <- adonis2(AvesPrim2 ~ Altitude + `Paved surface`+ Shrublands+ Ocean, data = Amb, by = "terms")

AICc.PERMANOVA2 <- function(adonis2.model) {

  # check to see if object is an adonis2 model...

  if (is.na(adonis2.model$SumOfSqs[1]))
    stop("object not output of adonis2 {vegan} ")

  RSS <- adonis2.model$SumOfSqs[ length(adonis2.model$SumOfSqs) - 1 ]
  MSE <- RSS / adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  nn <- adonis2.model$Df[ length(adonis2.model$Df) ] + 1

  k <- nn - adonis2.model$Df[ length(adonis2.model$Df) - 1 ]

  # AIC : 2*k + n*ln(RSS/n)
  # AICc: AIC + [2k(k+1)]/(n-k-1)

   AIC <- 2*k + nn*log(RSS/nn)
  AICc <- AIC + (2*k*(k + 1))/(nn - k - 1)

  output <- data.frame(AICc = AICc, k = k, N = nn)

  return(output)

}

DF<- data.frame(Modelo = paste0("Mod", 0:5), AICc = NA)

DF$AICc[1] <- AICc.PERMANOVA2(Mod0)$AICc
DF$AICc[2] <-  AICc.PERMANOVA2(Mod1)$AICc
DF$AICc[3] <- AICc.PERMANOVA2(Mod2)$AICc
DF$AICc[4] <-  AICc.PERMANOVA2(Mod3)$AICc
DF$AICc[5] <-  AICc.PERMANOVA2(Mod4)$AICc
DF$AICc[6] <-  AICc.PERMANOVA2(Mod5)$AICc

```

Permanova indicates that both winter and spring aassembles are determined by environmental variables such as altitude, ocean cover. Two variables are different for each season: during the winter, vegetation cover also determined assembles and during spring paved surfaces cover.

Both sampled seasons showed that few species contribute to differences among bird assembles found in the six habitats. 
Acccording to simper analysis, during winter eigth species showed to add to this difference between habitats. The contribution of each one of them go from 5 to 12%. These species are *Pelecanus thagus*, *Larus dominicanus*, *Coragyps atratus*, *Larosterna inca*, *Leucophaeus modestus*, *Columba livia*, *Turdus falcklandii*, and *Sephanoides sephanoides*. In each one of the paired habitat comparision the cumulative sum of two or more species reaches up to 29.3% of the explanation.

During the spring, six species show to add to the difference of bird assembles: *Leucophaeus pipixcan*, *Larus dominicanus*, *Columba livia*, *Larosterna inca*, *Phalacrocorax bougainvillii*, and *Pelecanus thagus*. The contribution of each species go from 5 to 11% to the difference, reaching 26.1% in the cumulative sum of two or more species contribution.
(For a complete species list of SIMPER, see Suplementary material SimperPrim, SimperInv).

```{r OccuInv, fig.width=10, fig.height=7}

library(sf)
library(raster)
library(ggrepel)
Pelecanus <-read_rds("PredInv_Pelecanus_thagus.rds")
Larus <-read_rds("PredInv_Larus_dominicanus.rds") 
Coragyps <-read_rds("PredInv_Coragyps_atratus.rds")
Larosterna <-read_rds("PredInv_Larosterna_inca.rds")
Leucophaeus <-read_rds("PredInv_Leucophaeus_modestus.rds")
Columba <-read_rds("PredInv_Columba_livia.rds")
Turdus <-read_rds("PredInv_Turdus_falcklandii.rds")
Sephanoides <-read_rds("PredInv_Sephanoides_sephaniodes.rds")

Nombres <- c("Pelecanus", "Larus", "Coragyps", "Larosterna", "Leucophaeus", "Columba", "Turdus", "Sephanoides")
Species <- list(Pelecanus, Larus, Coragyps, Larosterna, Leucophaeus, Columba, Turdus, Sephanoides)

temp <- list()

for(i in 1:length(Species)){
  temp[[i]] <- Species[[i]] %>% as("SpatialPixelsDataFrame") %>% as.data.frame()
  colnames(temp[[i]])[1] <- Nombres[i]
}

Species_DF <- temp %>% reduce(full_join) %>% dplyr::relocate(Pelecanus, .after = y) %>% pivot_longer(Pelecanus:Sephanoides, names_to = "Species", values_to = "Occupancy")

Sitios <- read.csv("/home/giorgia/Documents/Doctorado tesis/Monitoreo aves/MonitoreoVisualGit/Muestreo aves jun-jul 2019/Monitoreo_punto/Reg_completo_aves_inv.csv") %>% 
  dplyr::filter(Sitio %in% c("TERMINAL", "PLAYA LA BOCA", "PARQUE EL LITRE", "CALETA EL MEMBRILLO", "ROCA OCEANICA","SECTOR 5")) %>% 
  mutate(Sitio= str_replace(Sitio, "TERMINAL", "T"),
         Sitio= str_replace(Sitio, "PLAYA LA BOCA", "PLB"),
         Sitio= str_replace(Sitio, "PARQUE EL LITRE", "PEL"),
         Sitio= str_replace(Sitio, "CALETA EL MEMBRILLO", "CEM"),
         Sitio= str_replace(Sitio, "ROCA OCEANICA", "RO"),
         Sitio= str_replace(Sitio, "SECTOR 5", "PR")) %>% 
  dplyr::select(Sitio, Latitud, Longitud) %>% distinct() %>% st_as_sf(coords=c("Longitud","Latitud"),crs="+proj=longlat +datum=WGS84 +no_defs") %>% 
  st_transform(crs="+proj=utm +zone=19 +south +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")

OcuI <- ggplot() + geom_raster(data = Species_DF, aes(x = x, y = y, fill = Occupancy)) + geom_sf(data = Sitios, color="red") +
    ggrepel::geom_text_repel(data = Sitios,  aes(label = Sitio,  geometry = geometry),  stat = "sf_coordinates", force = 1, 
nudge_x=-5000, nudge_y = 1000)+ 
  theme_classic() + scale_fill_viridis_c("Occupancy", option = "plasma", direction = -1) + labs(x = NULL, y = NULL) + facet_wrap(~Species, nrow=1)+ theme(legend.position = "none")

```

```{r OccuPrim, fig.width=8, fig.height=7}
Columba <-read_rds("PredPrim_Columba_livia.rds")
Larosterna <-read_rds("PredPrim_Larosterna_inca.rds")
Larus <-read_rds("PredPrim_Larus_dominicanus.rds")
Leucophaeus <-read_rds("PredPrim_Leucophaeus_pipixcan.rds")
Phalacrocorax <-read_rds("PredPrim_Phalacrocorax_bougainvillii.rds")
Pelecanus <-read_rds("PredPrim_Pelecanus_thagus.rds")

Nombres <- c("Columba", "Larosterna", "Larus","Leucophaeus", "Phalacrocorax", "Pelecanus")
Species <- list(Columba, Larosterna, Larus, Leucophaeus, Phalacrocorax, Pelecanus)

temp <- list()

for(i in 1:length(Species)){
  temp[[i]] <- Species[[i]] %>% as("SpatialPixelsDataFrame") %>% as.data.frame()
  colnames(temp[[i]])[1] <- Nombres[i]
}

Species_DF <- temp %>% reduce(full_join) %>% dplyr::relocate(Columba, .after = y) %>% pivot_longer(Columba:Pelecanus, names_to = "Species", values_to = "Occupancy")

Sitios <- read.csv("/home/giorgia/Documents/Doctorado tesis/Monitoreo aves/MonitoreoVisualGit/Muestreo aves jun-jul 2019/Monitoreo_punto/Reg_completo_aves_inv.csv") %>% 
 dplyr::filter(Sitio %in% c("TERMINAL", "PLAYA LA BOCA", "PARQUE EL LITRE", "CALETA EL MEMBRILLO", "ROCA OCEANICA","SECTOR 5")) %>% 
  mutate(Sitio= str_replace(Sitio, "TERMINAL", "T"),
         Sitio= str_replace(Sitio, "PLAYA LA BOCA", "PLB"),
         Sitio= str_replace(Sitio, "PARQUE EL LITRE", "PEL"),
         Sitio= str_replace(Sitio, "CALETA EL MEMBRILLO", "CEM"),
         Sitio= str_replace(Sitio, "ROCA OCEANICA", "RO"),
         Sitio= str_replace(Sitio, "SECTOR 5", "PR")) %>% 
  dplyr::select(Sitio, Latitud, Longitud) %>% distinct() %>% st_as_sf(coords=c("Longitud","Latitud"),crs="+proj=longlat +datum=WGS84 +no_defs") %>% 
  st_transform(crs="+proj=utm +zone=19 +south +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")

OcuP <- ggplot() + geom_raster(data = Species_DF, aes(x = x, y = y, fill = Occupancy)) + geom_sf(data = Sitios, color="red") +
    ggrepel::geom_text_repel(data = Sitios,  aes(label = Sitio,  geometry = geometry),  stat = "sf_coordinates", force = 1, 
nudge_x=-5000, nudge_y = 1000)+ 
  theme_classic() + scale_fill_viridis_c("Occupancy", option = "plasma", direction = -1) + labs(x = NULL, y = NULL) + facet_wrap(~Species, nrow=1)

```

```{r Ocupancia, fig.width=13, fig.height=6}

layout <- "
AAAA
BBB#
"
OcuI + OcuP+plot_annotation(tag_levels = 'A')+ plot_layout(design = layout)

```

Predictions of ocurrence of those bird species on the studied area, indicates differences of habitat use (Fig. \@ref(fig:OccuInv) and \@ref(fig:OccuPrim)). 
Most of the predictions (nine of 15 proyections) show higher levels of occupancy in the coast than in the inland. 
In winter two species are strictly marine, three are marine and are able to use part of the terrestrial habitats, two are present in the whole stydy area, and one terrestrial and able to use part of the coast.
In spring, four species are strictly marine and two are present in the whole stydy area (one of them with some restrictiones in the coast and inland).

Only four species are present in both seasons. From those, two of species  showed the same use of habitats between winter and sprin (**Larus dominicanus** and **Pelecanus thagus**), while the other two showed differences between seasons (**Columba livia** and **Larosterna inca**).
*Larus dominicanus* presented the maximun probability of occupancy in the whole area both winter and spring. In this case, there was none covariable in the model that determined the occupancy. 
Occupancy of *Pelcanus thagus* is almos identical between both seasons. Projection shows the maximun probability of ocupancy in the coast and probabilities of zero in inland areas. The ocurrence of this species is determined by the altitude (negative effect) and in the winter ausence of native forest, grave and the presence of ocean surface, and in spring the presence of native forest, and the ausence of shrubs and sand.

During winter, community assemblages were explain by three environmental variables: coastal urban habitats, crops and tree plantations. In spring, variables were coastal urban habitats and native forest.

*Larosterna inca* shows similar ocurrence in both seasons, however during the spring it expand a little to the inland in some specific parts of the study area. The ocurrence is determine

*Columba livia* . During the winter its presence is focussed in certain coastal areas, defined by the presence of sand (mapa y modelo). During the spring, its presence is high in almost the whole study area with exception of some small areas. According to the model selected, the presence of native forest and sand, and the ausence of grave and tree plantation allow this pattern.

```{r}
#Resultados occupancia para especies dde Simper
ModelosOccuPrim <- read_rds("ModelosOccuPrim_Ms.rds")
#Para generar las ecuaciones **Resultado_OccuPrim_Ms.rds** y **Resultado_OccuInv_Ms.rds**
```

```{r Comportamiento, fig.height=8, fig.width=10}
# ComportamientoInv <- read_rds("ComportamientoInv.rds")
# I <-ggplot(ComportamientoInv) + 
#   geom_col(aes(x=AMBIENTE, y=Total_ind, fill=Comportamiento), color="black") + theme_classic()+ facet_wrap(~Cientifico, ncol=4, scales = "free_y")+  
#   theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))+ ylab("Número individuos") +xlab(" ")+
#   scale_fill_manual(aesthetics = "fill", values = c('#a6cee3','#1f78b4','#33a02c','#e31a1c','#fdbf6f','#ff7f00'), name="Bird behavior")
# 
# ComportamientoPrim <- read_rds( "ComportamientoPrim.rds")
# 
# P <-ggplot(ComportamientoPrim) + 
#   geom_col(aes(x=AMBIENTE, y=Total_ind, fill=Comportamiento), color="black") + theme_classic()+ facet_wrap(~Cientifico, ncol=3, scales = "free_y")+  
#   theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 1))+ ylab("Número individuos") +xlab(" ")+
#   scale_fill_manual(aesthetics = "fill", values = c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00'), name="Bird behavior")
# 
# layout <- "
# AAAAAAAA
# BBBBBBB#
# "
# I + P + plot_layout(design = layout) +plot_annotation(tag_levels = 'A')
#

```

### Discussion/Conclusion

1. Revising previous research- summaries
[Revisit previous research]
[Revisit Introduction to recall specific weakness in methodology used in previous studies]

2. Mapping 
[Revisit methodology used in this study]

This study gives information about not only bird species richness found in coastal urban habitats, also describe species identity in terms of assembles adapted in specific conditions within the city and bird ocurrence in continuos conditions

[Revisit and summaries the results]
Coastal urban environment studied present specific assembles of birds and they are determined mostly by variables such as altitude and ocean surface surrounding that areas. From the species ocurrence in those, there are strictly marine species and some can move to inland environments. Most of the terrestrial species are less strict and they manage to get closer to the ocean, specifically in areas with native vegetation such as natural beaches and natural rocky shore (with hay percentage of shrubs and native forrest).

[Show where and how the present work fits into the research "map" in this field]

3. Achievements/contributions- refining implications
[Recall an aspect of the results that represents a positive achievement or contribution of this work]

Environmental variability within coastal urban areas could allow the co-ocurrence of marine and terrestrial bird species, which is reflected in specific asembles in the six studied environments.

[Focus in the meaning and implications of the achievements in this work]

[Note if one of the achievement or contributions of the work is a novelty]
[Describe implications of the results, including possible applications]

4. Limitations-current and future work-applications
[Describe limitations which should direct future research]
High variability among studied environments plays a difficult rol 
This is considered at the momnet of occupancy analysis and analysis are made considering continuos data
It is possible that buffer area of influence should be specific for each species

[Suggest a specific area to be addressed in future work]

Clergeau, P., Jokim€aki, J. & Savard, J.-P. (2001). Are urban birdcommunities influenced by the bird diversity of adjacent landscapes?J. Appl. Ecol., 38, 1122–1134:
- Test the random dispersal-assembly hypothesis
- whatmakes a species tolerant to urbanisation is less obvious (Evanset al.2011). According to our analyses, this is primarily related  to  life  history  and  the  way  individuals  use  the resources
- Evidence is nonetheless clearer for the importance of lifehistory.  Life  history  integrates  key  evolved  adaptationsexpected to influence the colonisation-extinction dynamicsthat ultimately sustain biodiversity (Stearns 2000; Saetheret al.2005), and hence has long been considered central tounderstanding the varying responses of organisms to environ-mental changes

para que una especies se catalogue de tolerante o no tolerante a la urbanizacion, primero tiene q estar presente en los alrededores para ver si es posible el que entre o no a la ciudad.

especies en areas intervenidas y modificadas por la urbanizacion
disponibilidad de alimentacion y refugio que permiten la ocurrencia de especies (no necesariamente su reproduccion)
adaptabilidad de las especies para la busqueda de estos recursos (plasticidad en la dieta de palomas y gaviotas)

Much of the species loss is associated with a lack of appropriate adaptations of most species for exploiting resources and avoiding risks of the urban environments.

Some species with seasonal differences are related to migratory behaviour, or specific nesting habitat requierements (**Leucophaeus modestus** in @cikutovic1988gonadal, **Sephanoides sephanoides**).

### References